---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=TRUE}
library(ggplot2)
unzip('activity.zip')
activity_data <- read.csv('activity.csv')
steps_per_date <- tapply(activity_data$steps, activity_data$date, sum)
steps_per_date <- data.frame(date=names(steps_per_date), steps=as.vector(steps_per_date))
```


## What is mean total number of steps taken per day?
```{r, echo=TRUE, fig.width=20, fig.height=15}
p <- ggplot(data=steps_per_date[!is.na(steps_per_date$steps),], aes(x=steps)) 
p <- p + geom_histogram() 
p <- p + xlab("Date")
p <- p + ylab("Steps")
p <- p + ggtitle("Total Steps by Day")
p <- p + theme(
  axis.text.x = element_text(angle = 90, size=20),
  axis.text.y = element_text(size=20),
  axis.title = element_text(size=20,face="bold"),
  title = element_text(size=30,face="bold"))
print(p)
summary(steps_per_date$steps)[3:4]
```



## What is the average daily activity pattern?

```{r, echo=TRUE, fig.width=20, fig.height=15}
steps_per_interval <- data.frame(interval=activity_data$interval, steps=activity_data$steps)
steps_per_interval <- steps_per_interval[complete.cases(steps_per_interval),]
steps_per_interval <- tapply(steps_per_interval$steps, steps_per_interval$interval, mean)
steps_per_interval <- data.frame(interval=as.factor(as.integer(names(steps_per_interval))), steps=as.vector(steps_per_interval))
p <- ggplot(data=steps_per_interval, aes(x=interval, y=steps, group=1))
p <- p + stat_summary(fun.y=mean, geom="line")
p <- p + xlab("Interval")
p <- p + ylab("Average Steps")
p <- p + ggtitle("Average Steps by Interval")
p <- p + theme(
  axis.text.x = element_text(angle = 90, size=20),
  axis.text.y = element_text(size=20),
  axis.title = element_text(size=20,face="bold"),
  title = element_text(size=30,face="bold"))
p <- p + scale_x_discrete(breaks = function(x) {x[c(TRUE, rep(FALSE, 5))]})
print(p)
max_interval = steps_per_interval[steps_per_interval$steps==max(steps_per_interval$steps), ]
max_interval
```
The `r max_interval$interval` interval has the highest average number of steps with `r max_interval$steps` steps.

## Imputing missing values
```{r, echo=TRUE, fig.width=20, fig.height=15}
missing_count = sum(!complete.cases(activity_data))
missing_count

# Impute missing values using the mean for the given interval.
imputed_activity_data <- activity_data
imputed_activity_data[is.na(imputed_activity_data$steps),]$steps <- steps_per_interval[match(imputed_activity_data[is.na(imputed_activity_data$steps),]$interval, steps_per_interval$interval), 2]

imputed_steps_per_date <- tapply(imputed_activity_data$steps, imputed_activity_data$date, sum)
imputed_steps_per_date <- data.frame(date=names(imputed_steps_per_date), steps=as.vector(imputed_steps_per_date))

p <- ggplot(data=imputed_steps_per_date, aes(x=steps)) 
p <- p + geom_histogram() 
p <- p + xlab("Date")
p <- p + ylab("Steps")
p <- p + ggtitle("Total Steps by Day")
p <- p + theme(
  axis.text.x = element_text(angle = 90, size=20),
  axis.text.y = element_text(size=20),
  axis.title = element_text(size=20,face="bold"),
  title = element_text(size=30,face="bold"))
print(p)
summary(imputed_steps_per_date$steps)[3:4]

```

There are `r missing_count` rows with missing values in the dataset.


## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE, fig.width=20, fig.height=15}
imputed_activity_data$weekend_or_weekday <- as.factor(as.vector(sapply(weekdays(as.Date(imputed_activity_data$date)), function(x) {if (x %in% c('Saturday', 'Sunday')) "weekend" else "weekday"})))
p <- ggplot(data=imputed_activity_data, aes(x=interval, y=steps))
p <- p + stat_summary(fun.y=mean, geom="line")
p <- p + facet_grid(weekend_or_weekday~.)
p <- p + xlab("Interval")
p <- p + ylab("Average Steps")
p <- p + ggtitle("Average Steps by Interval")
p <- p + theme(
  axis.text.x = element_text(angle = 90, size=20),
  strip.text = element_text(angle = 90, size=20),
  axis.text.y = element_text(size=20),
  axis.title = element_text(size=20,face="bold"),
  title = element_text(size=30,face="bold"))

print(p)
```

